---
title: "Coronavirus Statistics In Colombia"
output: 
  flexdashboard::flex_dashboard:
    theme: default
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: embed
    # DESARROLLADORES:
    # Gian Marco Astori P. 2020114003
    # Juan Diaz Guerrero   2020114035
    # Johan s. Hawkins B.  2020114034
    
    # Grupo: 03
---

```{r setup, include=FALSE}
library(flexdashboard) # visualizaciones de datos relacionados como un tablero.
library(knitr) #Conversion a archivo plano
library(plyr) #Guardar tablas de datos.
library(ggplot2) #Generar las graficas.
library(scales) #Asignación datos a la estética.
library(readxl) #Para leer el excel
library(plotly) #Crea gráficos interactivos con calidad de publicación.
library(modeest) #Dato estadisco (MODA)
library(carData)
```

```{r}
#Se activa las varibles del archivo excel
Dat_Covid <- read.csv2("covid19_data_INS_Col.csv")
attach(Dat_Covid)
#-------------------------------------------------------------------------------------------------
#DATOS BASICOS
#Colocamos los de (confirmados, recuperados, fallecidos, infectados) en varables para utilizarlo despues en las tarjetas 
confirmados <- length(Dat_Covid$Caso)
recuperados <-length(Dat_Covid$Ubicacion[Ubicacion=="Recuperado"])
fallecidos <- length(Dat_Covid$Ubicacion[Ubicacion=="Fallecido"])
infectados <- confirmados - (recuperados + fallecidos)
#-------------------------------------------------------------------------------------------------
#fecha
Fechas <- (as.Date(Dat_Covid$Fecha_muerte, format = "%d/%m/%Y"))
Mes <- data.frame(table(format(Fechas, "%m")))
#-------------------------------------------------------------------------------------------------
#colores
azul_claro <- "#0abde3"
azul_oscuro <- "#2e86de"
rojo <-"#ff6b6b"
amarillo <- "#feca57"
Azul_menta <- "#59BFB3"
rosa <-"#ff9ff3"
verde_claro <- "#1dd1a1"
Morado <- "#5f27cd"
Azul_gris <- "#82B5C2"
Morado_claro <- "#be92eb"

#-------------------------------------------------------------------------------------------------
#TOTALES DE HOMBRES Y MUJERES
#sumamos la cantidad total de hombre (M,m) y lo guardamos en una misma variable (totalHombres)..asi mismo con las mujere
M1 <- length(Dat_Covid$Sexo[Sexo=="M"])
M2 <- length(Dat_Covid$Sexo[Sexo=="m"])
F1 <- length(Dat_Covid$Sexo[Sexo=="F"])
F2 <- length(Dat_Covid$Sexo[Sexo=="f"])
totalHombres <- M1+M2
totalMujeres <- F1+F2
#-------------------------------------------------------------------------------------------------
# Fechas
Fechas_not <- (as.Date(Dat_Covid$Fecha.Not, format = "%d/%m/%Y"))
Mes_not <- data.frame(table(format(Fechas_not, "%m")))

#-------------------------------------------------------------------------------------------------
#DEPARTAMENTO CON MAYOR CONTAGIO
#buscamos cual es el departamento mas afectado (depart) y la cantidad de contagio en ese departamento (casoDepart).. unimos esos datos en una misma variables (Dat_Depart) para mostrarlo en la tarjeta
depart <- mfv(Dat_Covid$nombre_depa)
casoDepart <- length(Dat_Covid$nombre_depa[nombre_depa==depart])
Dat_Depart <- paste(casoDepart, "cases in", depart)
#Santa Marta
Santa_Marta <- length(Dat_Covid$nombre_depa[nombre_depa=="Santa Marta D.T. y C."])

#-------------------------------------------------------------------------------------------------
#PER_ETN_2
#se declara una variable que guarda una funcion para poder utilizarlo mas facilmente en la tarjeta
x <- c("Indígena", "Negro", "Otro", "Rom", "")
var_per_ent_2 <- function(etnia2){length(Dat_Covid$per_etn_2[per_etn_2==etnia2])}

#-------------------------------------------------------------------------------------------------
#PAIS_VIAJO_1_NOM
#se organiza los paises por continentes en vectores y se le asigna a una variable
C_americano <- c("ARGENTINA", "BRASIL", "CANADÁ", "CHILE", "COSTA RICA", "CUBA", "ECUADOR", "ESTADOS UNIDOS", "ESTADOS UNIDOS DE AMÉRICA", "GUATEMALA", "HONDURAS", "JAMAICA", "MEXICO", "MÉXICO", "NICARAGUA", "PANAMA", "PANAMÁ", "PERU", "PERÚ", "PUERTO RICO", "REPÚBLICA DOMINICANA", "URUGUAY", "VENEZUELA"); Tamaño_America <- length(C_americano)

C_europeo <- c("ALEMANIA", "ANTILLAS NEERLANDESAS", "ARUBA", "BELGICA", "BOSNIA HERZEGOVINA", "CROACIA", "CURAZAO", "ESPAÑA", "FRANCIA", "GRECIA", "IRLANDA", "ITALIA", "PORTUGAL", "REINO UNIDO DE GRAN BRETAÑA E IRLANDA", "SUECIA", "SUIZA"); Tamaño_Europa <- length(C_europeo)

C_africa <- c("EGIPTO", "KENYA", "MARRUECOS"); Tamaño_Africa <- length(C_africa)

C_asia <- c("ARABIA SAUDÍ", "ARABIA SAUDITA", "BANGLADESH", "EMIRATOS ARABES UNIDOS", "ISRAEL", "REPÚBLICA ÁRABE SIRIA", "TURQUÍA"); Tamaño_Asia <- length(C_asia)

C_oceania <- c("AUSTRALIA"); Tamaño_Oceania <- length(C_oceania)

C_UNREGISTERED <- length(Dat_Covid$Pais_viajo_1_nom[Pais_viajo_1_nom==""])



#Calculamos cuantos viajante hay en cada continente utilizando ciclos (while), y guardando numero en la variable (total+nombredelcontinente)
#conteo C_americano

i1 <-1
totalAmerica <- 0
l1 <- 1
while(i1 <= Tamaño_America ){
  longitudAmerica <- length(Dat_Covid$Pais_viajo_1_nom[Pais_viajo_1_nom==C_americano[c(l1)]])
  totalAmerica <- longitudAmerica + totalAmerica 
  l1<- l1 + 1
  i1 <- i1 + 1
}

#conteo C_europeo
i2 <- 1
totalEuropa <- 0
l2 <- 1
while(i2 <= Tamaño_Europa ){
  longitudEuropa <- length(Dat_Covid$Pais_viajo_1_nom[Pais_viajo_1_nom==C_europeo[c(l2)]])
  totalEuropa <- longitudEuropa + totalEuropa
  l2 <- l2 + 1
  i2 <- i2 + 1
}

#conteo C_africa
i3 <- 1
totalAfrica <- 0
l3 <- 1
while(i3 <= Tamaño_Africa ){
  longitudAfrica <- length(Dat_Covid$Pais_viajo_1_nom[Pais_viajo_1_nom==C_africa[c(l3)]])
  totalAfrica <- longitudAfrica + totalAfrica
  l3 <- l3 + 1
  i3 <- i3 + 1
}

#conteo C_asia
i4 <- 1
totalAsia <- 0
l4 <- 1
while(i4 <= Tamaño_Asia ){
  longitudAsia <- length(Dat_Covid$Pais_viajo_1_nom[Pais_viajo_1_nom==C_asia[c(l4)]])
  totalAsia <- longitudAsia + totalAsia
  l4 <- l4 + 1
  i4 <- i4 + 1
}

#conteo C_oceania
i5 <- 1
totalOceania <- 0
l5 <- 1
while(i5 <= Tamaño_Oceania ){
  longitudOceania <- length(Dat_Covid$Pais_viajo_1_nom[Pais_viajo_1_nom==C_oceania[c(l5)]])
  totalOceania <- longitudOceania + totalOceania
  l5 <- l5 + 1
  i5 <- i5 + 1
}

#-------------------------------------------------------------------------------------------------
# Estado
Leve_1 <- length(Dat_Covid$Estado[Estado=="Leve"])
Moderado_2 <- length(Dat_Covid$Estado[Estado=="Moderado"])
Grave_3 <- length(Dat_Covid$Estado[Estado=="Grave"])
Asintomático_4 <- length(Dat_Covid$Estado[Estado=="Asintomático"])
Fallecido_5 <- length(Dat_Covid$Estado[Estado=="Fallecido"])
N_A_6 <- length(Dat_Covid$Estado[Estado=="N/A"])
```


Home
===================================================================

Row
-------------------------------------------------------------------

### CORONAVIRUS INFECTIONS COLOMBIA
```{r}
#Variable 1
valueBox(
  confirmados,
  color = azul_oscuro,
  icon = "fa-globe-americas",
)
```

### TOTAL DEATHS
```{r}
valueBox(
  fallecidos,
  color = rojo,
  icon = "fa-dizzy"
)
```

### CURRENTLY INFECTED
```{r}
valueBox(
  infectados,
  color = amarillo,
  icon = "fa-plus-square"
)
```

### TOTAL RECOVERED
```{r}
valueBox(
  recuperados,
  color = verde_claro,
 	icon = "fa-child"
)
```

Row
----------------------------------------------------------------------------------
   
### CONFIRMED MEN 
```{r}
#Variable 2 parte 1
valueBox(
  totalHombres,
  color = azul_claro,
 	icon = "fa-male"
)
```

### CONFIRMED WOMEN
```{r}
#Variable 2 parte 2
valueBox(
  totalMujeres,
  color = rosa,
  icon = "fa-female"
)
```

### MOST AFFECTED DEPARTMENT
```{r}
valueBox(
  Dat_Depart,
  color = Morado,
  icon = "fa-map-marker-alt",
)
```

### CASES IN SANTA MARTA D.T.C.H
```{r}
valueBox(
  Santa_Marta,
  color = Morado_claro,
  icon = "fa-map-marked-alt",
)
```

Row {.tabset}
--------------------------------------------------

### INDIGENOUS GROUPS
```{r}
valueBox(
  paste(var_per_ent_2(x[c(1)]),"Indigenous people were affected"),
  color = Azul_menta,
  icon = "fa-user-friends",
)
```

### AFROAMERICAN GROUPS
```{r}
valueBox(
  paste(var_per_ent_2(x[c(2)]),"Afroamerican people were affected"),
  color = Azul_menta,
  icon = "fa-user-friends",
)
```

### OTHER GROUPS
```{r}
valueBox(
  paste(var_per_ent_2(x[c(3)]),"Other people were affected"),
  color = Azul_menta,
  icon = "fa-user-friends",
)
```

### ROM GROUPS
```{r}
valueBox(
  paste(var_per_ent_2(x[c(4)]),"Rom people were affected"),
  color = Azul_menta,
  icon = "fa-user-friends",
)
```

### UNREGISTERED GROUPS
```{r}
valueBox(
  paste("",var_per_ent_2(x[c(5)]),"Unregistered people were affected"),
  color = Azul_menta,
  icon = "fa-user-friends",
)
```

Row {.tabset} 
--------------------------------------------------

### CASES - AMERICAN C.
```{r}
valueBox(
  paste(totalAmerica,"cases in America"),
  color = Azul_gris,
  icon = "fa-globe-americas"
)
```

### CASES - EUROPEAN C.
```{r}
valueBox(
  paste(totalEuropa,"cases in Europe"),
  color = Azul_gris,
  icon = "fa-globe-africa"
)
```

### CASES - AFRICAN C.
```{r}
valueBox(
  paste(totalAfrica,"cases in Africa"),
  color = Azul_gris,
  icon = "fa-globe-africa"
)
```

### CASES - ASIAN C.
```{r}
valueBox(
  paste(longitudAsia,"cases in Asia"),
  color = Azul_gris,
  icon = "fa-globe-asia"
)
```

### CASES - OCEANIC C.
```{r}
valueBox(
  paste(totalOceania,"cases in Oceania"),
  color = Azul_gris,
  icon = "fa-globe-asia"
)
```

### UNREGISTERED
```{r}
valueBox(
  paste(C_UNREGISTERED,"cases from unregistered countries"),
  color = Azul_gris,
  icon = "fa-globe"
)
```

Graphics
===================================================================

Row
--------------------------------------------------------------------

### F1 VARIATIONS OF COVID-19 CASES IN COLOMBIA
```{r}
plot_ly(
  x = c("Deceased", "Infected", "Recovered"),
  y = c(fallecidos, infectados, recuperados),
  color = ~c("Deceased", "Infected", "Recovered"),
  colors = c(rojo, amarillo, verde_claro),
  name = c("Deceased", "Infected", "Recovered"),
  type = "bar",
  height = 250, width = 320
)
```

### F2 COVID-19 CASES - FEMALE & MALE
```{r}
labels = c("MEN", "WOMEN")
values = c(totalHombres, totalMujeres)

plot_ly(
  type = "pie",
  labels = labels,
  values = values, 
  textinfo = "label+percent",
  insidetextorientation ="radial",
  height = 250, width = 320,
  marker = list(colors = c(azul_oscuro, rosa))
)
```

### F3 NUMBER OF DECEASED PER MONTH
```{r}
plot_ly(
  x = c("March", "April", "May", "June", "July", "August", "September"),
  y = c(29, 340, 898, 3138, 8124, 9459, 2710),
  color = ~c("March", "April", "May", "June", "July", "August", "September"),
  colors = c(rainbow(7)),
  name = c("March", "April", "May", "June", "July", "August", "September"),
  type = "bar",
  height = 250, width = 320
)
```

### F4 NOTIFIED DISEASED PER MONTH
```{r}
Months_of_notification<-Mes_not
Notified_people<-Mes_not$Freq
Specific_month<-Mes_not$Var1
fig<-plot_ly(
  data = Months_of_notification,
  Title = "Notified diseased per month",
  name = "Diseased",
  x = ~Specific_month,
  y = ~Notified_people,
  type = 'scatter',
  mode = 'lines',
  height = 250, width = 320
)

fig<- fig %>% layout(
    xaxis = list(title = "Specific months"),
    yaxis= list(title= "Notified diseased"),
    margin = list(l = 100)
)
fig
```

Row
--------------------------------------------------------------------

### F5 DISEASED INDIVIDUALS FROM DIFFERENT ETHNIC GROUPS
```{r}
labels = c("INDIGENOUS", "AFROAMERICAN", "OTHER", "ROM", "UNREGISTERED")
values = c(var_per_ent_2(x[c(1)]), var_per_ent_2(x[c(2)]), var_per_ent_2(x[c(3)]), var_per_ent_2(x[c(4)]), var_per_ent_2(x[c(5)]))

fig_2 <- plot_ly(
        type = "pie",
        labels = labels,
        values = values, 
        textinfo = "label+percent",
        insidetextorientation ="radial",
        height = 250, width = 320,
        textposition = 'inside',
        textinfo = 'label'
)

fig_2
```

### F6 AMOUNT OF DISEASED PER AGES
```{r}
Ages <- data.frame(table(Dat_Covid$Edad))
Amount <- Ages$Freq
Specific_Ages <- Ages$Var1

fig_3 <- plot_ly(
  data = Ages,
  x = ~Specific_Ages,
  y = ~Amount,
  type = 'scatter',
  mode = 'lines',
  height = 250, width = 320,
  color = I('red')
)

fig_3 <- fig_3 %>% layout(
    xaxis = list(title = "Specific ages"),
    yaxis= list(title= "Amount of diseased"),
    margin = list(l = 100)
)
fig_3
```

### F7 CASES IN REGISTERED CONTINENTS
```{r}
plot_ly(
  x = c("AMERICAN C.", "EUROPEAN C.", "AFRICAN C.", "ASIAN C.", "OCEANIC C."),
  y = c(totalAmerica, totalEuropa, totalAfrica, longitudAsia, totalOceania),
  color = ~c("AMERICAN C.", "EUROPEAN C.", "AFRICAN C.", "ASIAN C.", "OCEANIC C."),
  colors = c(rainbow(5)),
  name = c("AMERICAN C.", "EUROPEAN C.", "AFRICAN C.", "ASIAN C.", "OCEANIC C."),
  type = "bar",
  height = 250, width = 320
)
```

### F8 VARIETY OF CONDITIONS
```{r}
plot_ly(
  x = c("Light","Moderate","Severe","Asymptomatic","Deceased","N/A"),
  y = c(Leve_1, Moderado_2, Grave_3, Asintomático_4, Fallecido_5, N_A_6),
  color = ~c("Light","Moderate","Severe","Asymptomatic","Deceased","N/A"),
  colors = c(rainbow(6)),
  name = c("Light","Moderate","Severe","Asymptomatic","Deceased","N/A"),
  type = "bar",
  height = 250, width = 320
)
```

Conclusions
===================================================================

Row
-------------------------------------------------------------------

### Conclusions gathered from graphs and info cards 


- We can conclude that it's not a mortal disease, as only  `r round((fallecidos/confirmados*100),2)`% Of the infected individuals in Colombia have died.
 
- The point with the highest amount of diseased in Colombia is `r depart` with a total of  `r casoDepart` cases.
 
- The virus affects men the same way (`r round((totalHombres/confirmados*100),2)`%) as women (`r round((totalMujeres/confirmados*100),2)`%)
 
- In graph number 4, data has an ascending pattern that increases exponentially, reaching its peak in month 07 (septiembre) and 08 (agosto).

- To work with dates, we use as.date, format = %d/%m/%y.

- We have to use data.frame to put dates in graphs or cards.

- The population with the highest index of diseased individuals is the one that goes from 23 to 40 year olds.

- The continent with the highest amount of imported cases is the american one.

- The most affected ethnic group is the afroamerican one.

- The rate of recovered individuals (`r round((recuperados/confirmados*100),2)`%), is higher than the current rate of active infected individuals (as of the months reported in the files) (`r round((infectados/confirmados*100),2)`%) and the rate of deceased individuals (`r round((fallecidos/confirmados*100),2)`%).

